---
title: "Highest-rated dark chocolate bars by country"
author: "Oscar Amarilla"
date: "2023-08-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. The client

Chocolate and Tea is an up-and-coming chain of cafes combines an extensive menu of fine teas with chocolate bars from around the world. Their diverse selection includes everything from plantain milk chocolate, to tangerine white chocolate, to dark chocolate with pistachio and fig. The encyclopedic list of chocolate bars is the basis of Chocolate and Tea's brand appeal.

Chocolate bar sales are the main driver of revenue of Chocolate and Tea, so they aim to serve chocolate bars that are highly rated by professional critics. The management team regularly updates the chocolate bar list in order to align with the latest ratings and to ensure that the list contains bars from a variety of countries. They also continually adjust the menu to make sure it reflects the global diversity of chocolate production

## 2. The task

The management team of Chocolate and Tea asked to collect and analyze data on the latest chocolate ratings. In particular, they'd like to know which countries produce the highest-rated bars of super dark chocolate (a high percentage of cocoa).

## 3. Importing the necessary packages

The packages tidyverse, dplyr, ggplot2, and janitor are those required in this project.

```{r Importing the packages}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(janitor)
```

## 4. The dataset

The dataset received is a csv file that features the latest ratings for thousands of chocolates from around the world.

### 4.1 Importing the data

```{r Importing the dataset}
cocoa_df <- read_csv("cacao.csv")
head(cocoa_df)
```

### 4.2 General information of the dataset

The features of the dataset are the following

```{r}
colnames(cocoa_df)
```

The columns are defined as:

-   **Company (Maker-if known):** Name of the company manufacturing the bar.
-   **Specific Bean Origin or Bar Name:** The specific geo-region of origin for the bar.
-   **REF:** A value linked to when the review was entered in the database. Higher = more recent.
-   **Review Date:** Date of publication of the review.
-   **Cocoa Percent:** Cocoa percentage (darkness) of the chocolate bar being reviewed.
-   **Company Location:** Manufacturer base country.
-   **Rating:** Expert rating for the bar.
-   **Bean Type:** The variety (breed) of bean used, if provided.
-   **Broad Bean Origin:** The broad geo-region of origin for the bean.

Two of the most relevant variables for the business goal are the Cocoa Percent and the Rating. With respect to the Rating, the value scale is

-   5= Elite (Transcending beyond the ordinary limits)
-   4= Premium (Superior flavor development, character and style)
-   3= Satisfactory(3.0) to praiseworthy(3.75) (well made with special qualities)
-   2= Disappointing (Passable but contains at least one significant flaw)
-   1= Unpleasant (mostly unpalatable)

```{r}
str(cocoa_df)
```

### 4.3 Data cleaning and organization

The first thing is to change the name of some columns because they are in an inconvenient format.

```{r}
cocoa_df <- cocoa_df %>% # This method doesn't works with the first row.
  rename(Bean_origin_or_bar_name="Specific Bean Origin\nor Bar Name", Review_date="Review\nDate",
Cocoa_percent="Cocoa\nPercent", Location="Company\nLocation", Bean_type="Bean\nType", Broad_bean_origin="Broad Bean\nOrigin")
cocoa_df <- cocoa_df %>%
    rename_with(~ "Company", 1)
colnames(cocoa_df)
```

The next step is to look for null and other spurious values (empty, incoherent data, characters instead numbers or vice versa, etc.).

```{r}
    null_values <- is.na(cocoa_df)
    col_null_counts <- colSums(null_values)
    col_null_counts
```

The stakeholder has established that most relevant variables for the analysis are Rating, Cocoa_percent, and Location. Therefore, the missing values in the Bean_type and Broad_bean_origin are not relevant for the analysis, so the analysis has to focus on this variables.

```{r}
unique(cocoa_df$Cocoa_percent)
```

```{r}
unique(cocoa_df$Location)
```

```{r}
unique(cocoa_df$Rating)
```

As can be seen, none of these columns contain spurious values, so the analysis can go on. A new data frame is established in order to reduce the number of features being manipulated. Before that, has been notice that the Cocoa_percent values are strings, so these values will be converted to float.

```{r}
cocoa_df$Cocoa_percent <- as.numeric(sub("%", "", cocoa_df$Cocoa_percent))
```

```{r}
trimmed_cocoa_df <- cocoa_df %>%
  select(Company, Rating, Cocoa_percent, Location)
colnames(trimmed_cocoa_df)
```

### 4.4 Exploratory analysis

The rating distribution will show range of quality of the cocoa bars considered in the dataset.

```{r}
ggplot(data = trimmed_cocoa_df, aes(x = Rating)) +
 # geom_density(color = "red", size = 1) +
  geom_histogram(fill = "steelblue", color = "black") +
  labs(title = "Distribution of the rating scores", x = "Rating", y = "Frequency")
```

The products offered are mainly, with respect to the rating, disappointing to premium, being predominately the presence of 3.5 score of rating.

```{r}
ggplot(data = trimmed_cocoa_df, aes(x = Cocoa_percent)) +
 # geom_density(color = "red", size = 1) +
  geom_histogram(fill = "steelblue", color = "black", stat = "count") +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Distribution of the cocoa percentage", x = "Percentage", y = "Frequency")
```

The distribution of the percent of cocoa in the bars doesn't has a regular pattern although there is a concentration of products between the 60% and 80%. The 70% of cocoa chocolate bars are the most common in the dataset, there is no product with less than 40% of cocoa in their recipe.

Looking for correlation between the cocoa percentage and the rating, a plot relating this two variables is shown.

```{r}
ggplot(data = cocoa_df) + geom_point(mapping = aes(x = Cocoa_percent, y = Rating))
```

There is correlation between this two attributes of the chocolate bars. But what is visible in this plot is that there is a high density area is the box determined by the box of 2.5 and 4 rating score, and the 60% and 80% of cocoa percent.

Another interesting feature to visualize is the location of the manufacturing.

```{r}
ggplot(data = trimmed_cocoa_df, aes(x = Location)) +
  geom_bar(fill = "steelblue") +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Bar plot of locations of manufacturing", x = "Locations", y = "Count")
```

The biggest producer of cocoa bar is the U.S.A., followed by France, then Canada, and in the fourth place is the U.K..

After this brief analysis, the conclusion is to narrow down the offers to those bars with a rating equal to or higher than 3.75 points. It has to be taken into account that Chocolate and Tea, the stakeholders, consider a bar to be super dark chocolate if the bar's cocoa percentage is greater than or equal to 80%. Therefore, it was decided to create a new data frame that groups those chocolate bars that meet these two conditions.

```{r}
potential_suppliers <- trimmed_cocoa_df %>%
  filter(Rating >= 3.75 & Cocoa_percent >= 80)
potential_suppliers
```

Finally, the remaining list of potential suppliers has a mean rating score of 3.8 with a standard deviation of 0.1. The mean cocoa percentage in the bar recipes is 82.5% with a standard deviation of 4.1. However, it is important to note that this last value may not be representative due to the presence of products with 88% and 90% cocoa. Therefore, a more accurate measure to describe this variable is to look at the median value, which is 80% cocoa in the bars.

```{r}
potential_suppliers %>%
  summarise(mean(Rating),sd(Rating), mean(Cocoa_percent), sd(Cocoa_percent), median(Cocoa_percent))
```

Most of the products offered came from the U.S.A., but the highest quality chocolate bars are manufactured at France and Canada.

```{r bar_rated_by_country, echo=FALSE}
ggplot(data = potential_suppliers) +
  geom_bar(mapping = aes(x = Location, fill=Rating))
```

Finally, the distribution of cocoa percentage of the resulting list shows that companies at Canada and the U.S.A. are those who offer the chocolate bars with highest presence of cocoa in their recipe.

```{r facet_plot, echo=FALSE}
ggplot(data = potential_suppliers) +
  geom_bar(mapping = aes(x = Cocoa_percent)) +
  facet_wrap(~Location)
```

## 4. Conclusions

Through this analysis, the list of 1795 potential suppliers was successfully narrowed down to just eight. This reduction was based on a minimum rating score of 3.75 out of 5 and a minimum cocoa percentage of 80%, which is considered as super dark chocolate according to the stakeholders' criteria. The suggestion is to focus on two of these eight suppliers: Pralus, a French company, and Soma, a Canadian company. This recommendation is supported by the fact that both companies have received a score of 4 out of 5 from the experts, indicating that their products exhibit premium quality, characterized by superior flavor development, character, and style.
